\HeaderA{bg}{Set up Background}{bg}
\aliasA{bg3d}{bg}{bg3d}
\aliasA{rgl.bg}{bg}{rgl.bg}
\keyword{dynamic}{bg}
\begin{Description}\relax
Setup the background environment of the scene.
\end{Description}
\begin{Usage}
\begin{verbatim}
bg3d(...) 
rgl.bg( sphere = FALSE, fogtype = "none", color=c("black","white"), 
        back="lines", ...)

\end{verbatim}
\end{Usage}
\begin{Arguments}
\begin{ldescription}
\item[\code{fogtype}] fog type:
\describe{
\item["none"] no fog
\item["linear"] linear fog function
\item["exp"] exponential fog function
\item["exp2"] squared exponential fog function
}

\item[\code{sphere}] logical, if true, an environmental sphere geometry is used for the background decoration.

\item[\code{color}] Primary color is used for background clearing and as fog color.
Secondary color is used for background sphere geometry. See \code{\LinkA{rgl.material}{rgl.material}} for details.

\item[\code{back}] Specifies the fill style of the sphere geometry. See \code{\LinkA{rgl.material}{rgl.material}} for details.

\item[\code{ ... }] Material properties. See \code{\LinkA{rgl.material}{rgl.material}} for details.
\end{ldescription}
\end{Arguments}
\begin{Details}\relax
If sphere is set to TRUE, an environmental sphere enclosing the whole scene is drawn.
\end{Details}
\begin{SeeAlso}\relax
\code{\LinkA{rgl.material}{rgl.material}}
\end{SeeAlso}
\begin{Examples}
\begin{ExampleCode}
  rgl.open()
  
  # a simple white background
  
  bg3d("white")

  # the holo-globe (inspired by star trek):

  rgl.bg(sphere=TRUE, color=c("black","green"), lit=FALSE, back="lines" )

  # an environmental sphere with a nice texture.

  rgl.bg(sphere=TRUE, texture=system.file("textures/sunsleep.png", package="rgl"), 
         back="filled" )
\end{ExampleCode}
\end{Examples}

