\HeaderA{texts}{add text}{texts}
\aliasA{rgl.texts}{texts}{rgl.texts}
\aliasA{text3d}{texts}{text3d}
\aliasA{texts3d}{texts}{texts3d}
\keyword{dynamic}{texts}
\begin{Description}\relax
Adds text to the scene. The text is positioned in 3D space. 
A bitmap font which is always oriented towards the camera is used.
\end{Description}
\begin{Usage}
\begin{verbatim}
rgl.texts(x, y = NULL, z = NULL, text, adj = 0.5, justify, ...)
text3d(x, y = NULL, z = NULL, texts, adj = 0.5, justify, ...) 
texts3d(x, y = NULL, z = NULL, texts, adj = 0.5, justify, ...) 
\end{verbatim}
\end{Usage}
\begin{Arguments}
\begin{ldescription}
\item[\code{x, y, z}] point coordinates.  Any reasonable way of defining the
coordinates is acceptable.  See the function \code{\LinkA{xyz.coords}{xyz.coords}}
for details.
\item[\code{text}] text character vector to draw
\item[\code{texts}] text character vector to draw
\item[\code{adj}] one value specifying the horizontal adjustment 
\item[\code{justify}] (deprecated, please use adj instead) character string
specifying the horizontal adjustment; options are \code{"left"}, \code{"right"},
\code{"center"}.

\item[\code{ ... }] Material properties. See \code{\LinkA{rgl.material}{rgl.material}} for details.
\end{ldescription}
\end{Arguments}
\begin{Details}\relax
The \code{adj} parameter determines the position of the text relative to the 
specified coordinate.  Use \code{adj = 0} to place the left bottom corner at
\code{(x,y,z)}, \code{adj = 0.5} to center the text there, and \code{adj = 1}
to put the right bottom corner there.

\code{text3d} and \code{texts3d} draw text using the \LinkA{r3d}{r3d}
conventions. These are synonyms; the former is singular to be
consistent with the classic 2-D graphics functions, and the latter is
plural to be consistent with all the other graphics primitives.  Take
your choice!

If any coordinate or text is \code{NA}, that text is not plotted.
\end{Details}
\begin{Value}
The object ID of the text object is returned invisibly.
\end{Value}
\begin{SeeAlso}\relax
\code{\LinkA{r3d}{r3d}}
\end{SeeAlso}
\begin{Examples}
\begin{ExampleCode}
open3d()
text3d(rnorm(10)*100,rnorm(10)*100,rnorm(10)*100,text=1:10,adj = 0.5, 
       color=heat.colors(10))
\end{ExampleCode}
\end{Examples}

